### 2020-06-01 writeup/experiments.tex
intro

We connect our theory to pratical problems of interest.
First, we validate the single-task based metric on text and image classification tasks.
We show that for both datasets, single-task learning results can help predict positive or negative transfer in multi-task learning.
Second, \todo{}.
Third, we measure the data efficiency ratio of multi-task learning on six sentiment analysis tasks.
Finally, we validate the three components of our theory in Section \ref{sec_insight} on text classification tasks.

>>>>

We connect our theory to pratical problems of interest.
First, we validate the single-task based metric on text and image classification tasks.
We show that for both datasets, single-task learning results can help predict positive or negative transfer in multi-task learning.
Second, we show that our proposed incremental training schedule improves the training efficiency of multi-task learning for predicting a particular target task.
%Third, we measure the data efficiency ratio of multi-task learning on six sentiment analysis tasks.
Finally, we validate the three components of our theory and measure the data efficiency of multi-task learning on text classification tasks.

### 2020-06-01 writeup/intro.tex
intro

\begin{abstract}
When does multi-task learning outperform single-task learning?
In this work, we address this question by studying the performance of predicting a particular task given multiple tasks.
We consider a multi-task learning architecture that uses a shared layer for all tasks and a separate output layer for each task.
First, for the setting of high-dimensional linear regression, we provide a sharp analysis of the bias-variance tradeoff of multi-task learning estimators.
We develop a technical tool on the inverse of the sum of two random matrices that naturally arises from the analysis.
We extend the result to the related setting of transfer learning.
Second, we apply our tool to rigorously explain when multi-task learning outperforms single-task learning.
%And we theoretically analyze the benefit of multi-task learning for reducing the amount of labeled data needed to achieve comparable performance to single-task learning. %, which has been a key empirical finding in recent work.
Finally, we show practical implications of our theory for detecting and mitigating negative effects in image and text classification tasks.
\end{abstract}

@@

Our second contribution is to use our newly developed tool to rigorously explain negative transfer in multi-task learning.
In Section \ref{sec_insight}, we consider three factors including task model, data size and covariate shift.
\squishlist
		\item First, we define task dissimilarity as the distance between $\beta_1$ and $\beta_2$ using a simplified isotropic setting of two tasks.
		Using our technical tool, we show a sharp transition from positive to negative transfer as task dissimilarity increases.
		Furthermore, we show that negative transfer is more likely to occur when the source task labels are particularly noisy.
%		In Section \ref{sec_validate}, we validate the observation on text and image classification tasks.
%	In , we provide the trade-off between $\norm{\beta_1 - \beta_2}^2$ and a certain function $\Phi(\rho_1, \rho_2)$ to determine the type of transfer.
		\item Second, we consider data size.
	We begin by observing that adding more labeled data from the source task does not always improve performance.
	Then, we analyze the benefit of multi-task learning for reducing the amount of labeled data needed to achieve comparable performance to single-task learning, which is a key empirical finding of Taskonomy \cite{ZSSGM18}.
%	We define the \textit{data efficiency ratio} as the smallest $x$ such that if we only use an $$ fraction of labeled data, then the test error of $\hat{\beta}_t^{\MTL}$ matches that of the $\hat{\beta}_t^{\STL}$ on the entire set.
	%In Proposition \ref{prop_data_efficiency}, we show that the data efficiency ratio of an illustrative example is at most $\alert{\frac {1}{2\rho_t}}$.
	%In Section \ref{sec_validate}, we validate that performing multi-task learning can reduce the need for labeled data on 6 sentiment analysis tasks. %, we observe that just by using \alert{40\%} of the labeled data, the overall accuracy of multi-task learning matches that of learning every task in isolation.
		\item Finally, we show that covariate shift, measured by $\Sigma_1^{1/2}\Sigma_2^{-1/2}$, is another cause for suboptimal performance for $\hat{\beta}_t^{\MTL}$.
		We show that as $n_1 / n_2$ becomes large, the source task should have the same covariance matrix as the target task to ensure the best performance.
%		On the other hand, when $n_1 / n_2$ is small, there are counter examples where having the same covariance matrix is not necessarily the optimal choice.
\squishend

Our final contribution is to connect our theory to practical problems of interest.
First, we provide a metric to determine positive versus negative transfer by comparing the test accuracies of single-task models.
Second, \todo{}
Finally, we provide a new insight on the covariance alignment algorithm, proposed in \cite{WZR20}.
We show that as the size of the source task becomes larger compared to the target task, the covariance alignment algorithm produces more significant results.

>>>>

\begin{abstract}
When does multi-task learning outperform single-task learning?
In this work, we address this question by studying the performance of predicting a particular task given multiple tasks.
We consider a multi-task learning architecture that uses a shared layer for all tasks and a separate output layer for each task.
First, for the setting of high-dimensional linear regression, we provide a sharp analysis of the bias-variance tradeoff of multi-task learning estimators.
We develop a technical tool on the inverse of the sum of two random matrices that naturally arises from the analysis.
We extend the result to the related setting of transfer learning.
Second, we apply our tool to rigorously explain when multi-task learning outperforms single-task learning.
%And we theoretically analyze the benefit of multi-task learning for reducing the amount of labeled data needed to achieve comparable performance to single-task learning. %, which has been a key empirical finding in recent work.
Finally, we show practical implications of our theory for detecting and mitigating negative effects.
	For example, we propose an incremental training scheduler for improving the efficiency of multi-task training for predicting a particular task.
	We validate our theory and proposed method on image and text classification tasks.
\end{abstract}

@@

Our second contribution is to use our newly developed tool to rigorously explain negative transfer in multi-task learning.
In Section \ref{sec_insight}, we consider three components including task model, data size and covariate shift for a simplified isotropic setting of two tasks.
\squishlist
		\item We measure task dissimilarity by how small is the distance between $\beta_1$ and $\beta_2$.
		Using our technical tool, we explain a transition from positive to negative transfer as task similarity decreases.
%		Furthermore, we show that negative transfer is more likely to occur when the source task labels are particularly noisy.
%		In Section \ref{sec_validate}, we validate the observation on text and image classification tasks.
%	In , we provide the trade-off between $\norm{\beta_1 - \beta_2}^2$ and a certain function $\Phi(\rho_1, \rho_2)$ to determine the type of transfer.
		\item We show that increasing the data size of the source task does not always improve performance for the target task in multi-task learning.
	Then, we analyze the benefit of multi-task learning for reducing the amount of labeled data needed to achieve comparable performance to single-task learning, which is a key empirical finding of Taskonomy \cite{ZSSGM18}.
%	We define the \textit{data efficiency ratio} as the smallest $x$ such that if we only use an $$ fraction of labeled data, then the test error of $\hat{\beta}_t^{\MTL}$ matches that of the $\hat{\beta}_t^{\STL}$ on the entire set.
	%In Proposition \ref{prop_data_efficiency}, we show that the data efficiency ratio of an illustrative example is at most $\alert{\frac {1}{2\rho_t}}$.
	%In Section \ref{sec_validate}, we validate that performing multi-task learning can reduce the need for labeled data on 6 sentiment analysis tasks. %, we observe that just by using \alert{40\%} of the labeled data, the overall accuracy of multi-task learning matches that of learning every task in isolation.
		\item Finally, we show that covariate shift, measured by $\Sigma_1^{1/2}\Sigma_2^{-1/2}$, is another cause for suboptimal performance for $\hat{\beta}_t^{\MTL}$.
		We show that as $n_1 / n_2$ becomes large, having no covariate shift between the source and target tasks yields the optimal performance for the target task.
%		On the other hand, when $n_1 / n_2$ is small, there are counter examples where having the same covariance matrix is not necessarily the optimal choice.
\squishend

Our final contribution is to connect our theory to practical problems of interest.
First, we provide a single-task based metric to predict positive or negative transfer in multi-task learning.
While it is generally not well-understood when multi-task learning provides positive transfer, we show that the STL results are indicative of and can help understand MTL results on ChestX-ray14 \cite{chexnet17} and sentiment analysis datasets \cite{LZWDA18}.
Second, we design an incremental training schedule to improve the efficiency of multi-task training for predicting a particular task.
We show that our training schedule reduces the computational cost by $55\%$ compared to baseline multi-task training on the sentiment analysis dataset, while keeping the accuracy the same.
Finally, we validate the three components of our theory in Section \ref{sec_insight} and measure the data efficiency of multi-task learning on the sentiment analysis dataset.
